## Homework 4

<!--
Remember you can use the example4.Rmd or other files to find examples
of the codes you'll need to use to complete this assignment. I strongly
urge you to copy and paste relevant code that you know works from my
examples or instructions and then edit that code so you're working
with the appropriate variables for the homework.
-->

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

<!--
Read in the homework4.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
-->
First, I read in the data.
```{r}
dat<-read.csv("homework4.csv")
```


<!--
Check the structure of the data.
-->
Then I can check the structure of the data.
```{r}
str(dat)
```

The structure of the data has two variables, each of which has two factors.
The first variable is Condition, with the factors control and experimental.
The second variable is Agreement, with the factors no and yes. 

<!--
Create a table that shows how many people fell into each category.
-->
After the data is read in, I am able to create a table to show how many people are in each category.
```{r}
table(dat$condition, dat$agreement)
```

From the table, it appears that there are more yes answers for the experimental group.
This might indicate that the experimental condition was related to yes answers.
The control group did not differ between yes and no answers. 


<!--
Perform a chi-square test to determine whether or not the experimental
conditions and agreement with a statement are independent or related.
Also, interpret the results of your test and provide a a statistical
statement that supports your interpretation. As there's no way to
insert the chi-square symbol, you can just use chi-square instead
(see example4.Rmd).
-->
A chi-square test can now be performed to determine if experimental condition and agreement are independent.
```{r}
chisq.test(dat$condition,dat$agreement)
```

The results were Chi-square(1) = 1.2788, p = 0.2581 which reveals that experimental condition and agreement are in fact independant and not related despite there being a higher number of yes responses in the experimental condition.
