---
output: word_document
---
## Homework 5

<!--
Set the working directory.
No Rchunk is needed to accomplish this.
-->

I start by setting the working directory to the source file location.

<!--
Read in the homework5.csv data. For each of the following steps you
should preface your R commands with some plain English explanation.
The data are fabricated and include the number of words per minute
that people can type when listening to no music and when listening
to heavy metal.
-->

My first step is to read in the data.
```{r}
dat<-read.csv("homework5.csv")
```

<!--
Check the structure of the data.
-->
Then I can check the structure of the data.
```{r}
str(dat)
```

By checking the structure I can see that there are two variables. The first is the independant variable of music, which has two levels: metal and none. The second is the dependent variable of words per minute.

<!--
Find the mean words per minute for the two groups.
-->
Now I can find the mean words per minute for the two levels of the IV music.
```{r}
mean(dat$wpm[dat$music=="metal"])
mean(dat$wpm[dat$music=="none"])
```

This reveals that the mean words per minute for the condition music:metal is 48.04, and the mean words per minute for the condition music:none is 42.45.
The higher mean word per minute for the music:metal condition seems to indicate that participants in the condition music:metal typed more quickly than those in the music:none condition.

<!--
Run an independent t-test to determine if the mean wpm for the
no music group is different from the mean wpm from heavy metal
group. Interpret the results of your analysis after you run it.
-->
I can now run an independent t-test to find out if there actually is a difference in mean words per minute between the two levels of music metal and none.

```{r}
t.test(dat$wpm[dat$music=="metal"],
       dat$wpm[dat$music=="none"],
       var.equal=TRUE)
```

The results of the t-test reveal that there is a significant difference between the two conditions. Words per minute in the condition music:metal was higher than in the condition music:none t(62) = 2.33, p = .02298.

<!--
Run another independent t-test with the alternative hypothesis that
the heavy metal group types faster than the no music group. Interpret
the results of your analyses.
-->
I can also run an independent t-test to determine if the alternative hypothesis is true: that participants in the condition music:metal type faster than the participants in the condition music:none.
```{r}
t.test(dat$wpm[dat$music=="metal"],
       dat$wpm[dat$music=="none"],
       var.equal=TRUE,
       alternative="greater")
```
The results of the test confirm that participants in the condition music:metal type faster than participants in the condition music:none t(62) = 2.33, p = .01149

<!--
Create a bar graph of the data.
-->
I can create a bar graph to display the results by loading the ggplot2, gplots, and dplyr libraries.

```{r}
library(ggplot2)
library(gplots)
library(dplyr)
```

Then I can use dplyr to analyze the data. I can find the means and standard errors of the means and I can summarize the data for the different groups.
```{r}
temp<-dat%>%group_by(music)%>%
  summarize(means=mean(wpm), sems=sd(wpm)/sqrt(length(wpm)))
```
I am able to create a figure from this summarized data.
```{r}
f<-ggplot(temp,aes(x=factor(music),y=means))+
  geom_bar(stat="identity", color="black",
           fill=c("deepskyblue2", "deeppink"))+
  geom_errorbar(aes(ymax=means+sems,
                    ymin=means-sems),
                width=.2)+
  ggtitle("Words per Minute for Music Conditions")+
  labs(x="Music", y="Words per Minute (wpm)")+
  scale_x_discrete(breaks=c("metal","none"),
                   labels=c("Metal","None"))+
  theme(plot.title=element_text(size=15,
                                face="bold",
                                vjust=.5))+
  theme(axis.title.x=element_text(size=12,
                                  face="bold",
                                  vjust=-.25))+
  theme(axis.title.y=element_text(size=12,
                                  face="bold",
                                  vjust=.25))+
  theme(axis.text.x=element_text(size=10,
                                 face="bold",
                                 color="black"))+
  theme(axis.text.y=element_text(size=10,
                                 face="bold",
                                 color="black"))+
  coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),
                         max(temp$means)+2*max(temp$sems)))+
  theme(panel.border=element_blank(),
        axis.line=element_line())+
  theme(panel.grid.major.x=element_blank())+
  theme(panel.grid.major.y=element_line(color="darkgrey"))+
  theme(panel.grid.minor.y=element_blank())
f
```
<!--
Conduct and interpret a two-tailed paired t-test on the data.
-->
Another test I am able to perform on the data is a two-tailed paired t-test.
```{r}
t.test(dat$wpm[dat$music=="metal"],
       dat$wpm[dat$music=="none"],
       var.equal=TRUE,
       paired=TRUE)
```
The results reveal that those in the music:metal condition typed more quickly than those in the music:none condition t(31) = 12.153, p < .05.